---
title: "Plotting Trees with colorsk"
output: html_notebook
---

```{r setup, include=FALSE}

library(knitr)
library(tidyverse)
library(treeio)
library(readxl)
library(ggtree)
workdir="/atium/Data/Nanopore/cpowgs/trees/200630_tree/"
opts_knit$set(root.dir=workdir)

#Pass to bash
Sys.setenv(workdir=workdir)

```

Ok - the goal for this notebook is to generate plots of the raxml trees  we generated with colored tips, etc from the metadata from Trish

First read the meta info

```{r}

meta.klpn=read_excel("200630_trish_meta.xlsx", sheet="KLPN") %>%
  mutate(label=paste0("KLPN_", `Study ID`)) %>%
  select(label, MLST, Year) %>%
  replace_na(list(MLST="unknown", Year="unknown"))
meta.encl=read_excel("200630_trish_meta.xlsx", sheet="ENCL") %>%
  mutate(label=paste0("ENCL_", `STOCK ID`)) 
#  select(label, MLST, Year) %>%
#  replace_na(list(MLST="unknown", Year="unknown"))    
meta.esco=read_excel("200630_trish_meta.xlsx", sheet="ESCO") %>%
  mutate(label=paste0("ESCO_", `STUDY ID`)) %>%
  select(label, MLST, Year) %>%
  replace_na(list(MLST="unknown", Year="unknown"))
  
  
```

Now let's try to get the trees in from the raxml

```{r}

tree.root="/atium/Data/Nanopore/cpowgs/trees/200408_timp_tree"
tree.klpn=read.tree(file.path(tree.root, "raxml_klpn/RAxML_bestTree.testrapid.tree" ))
tbl.klpn=as_tibble(tree.klpn) %>%
  left_join(meta.klpn)
phy.klpn=as.treedata(tbl.klpn)

tree.encl=read.tree(file.path(tree.root, "raxml_encl/RAxML_bestTree.testrapid.tree" ))
tbl.encl=as_tibble(tree.encl) %>%
  left_join(meta.encl)
phy.encl=as.treedata(tbl.encl)

tree.esco=read.tree(file.path(tree.root, "raxml_esco/RAxML_bestTree.testrapid.tree" ))
tbl.esco=as_tibble(tree.esco) %>%
  left_join(meta.esco)
phy.esco=as.treedata(tbl.esco)



#phy.klpn=tree.klpn %<+% meta.klpn

```


Now let's plot the tree



```{r}
#pdf("~/Dropbox/timplab_data/cpowgs/200707_tree/klpn_tree.pdf")
print(ggtree(phy.klpn)+geom_tippoint(aes(color=as.factor(Year)))+theme(legend.position="right")+ggtitle("klpn"))
print(ggtree(phy.esco)+geom_tippoint(aes(color=as.factor(Year)))+theme(legend.position="right")+ggtitle("esco"))
print(ggtree(phy.encl)+geom_tippoint(aes(color=as.factor(Year)))+theme(legend.position="right")+ggtitle("encl"))
#dev.off()
```

Ok - now let's try to add MLST

Notes from: 
https://groups.google.com/forum/?utm_medium=email&utm_source=footer#!msg/bioc-ggtree/Q4LnwoTf1DM/uqYdYB_VBAAJ
https://yulab-smu.github.io/treedata-book/chapter2.html#data-integration

for groupOTU which is what I think I need, I'm supposed to crate a list of OTUs of the child node labels? meta.klpn has this info, right?  So what I need to do is somehow cluster

```{r}

grouped=meta.klpn %>%
  group_by(MLST) %>%
  summarize(mlst_group=list(label))

mlst_list=grouped$mlst_group
names(mlst_list)=grouped$MLST

phy.klpn.group=groupOTU(phy.klpn, mlst_list)

```

```{r}

pdf("~/Dropbox/timplab_data/cpowgs/200707_tree/klpn_tree_mlst.pdf")
print(ggtree(phy.klpn)+geom_tippoint(aes(color=as.factor(MLST)))+theme(legend.position="right")+ggtitle("klpn"))
print(ggtree(phy.klpn, aes(color=as.factor(MLST)))+theme(legend.position="right")+ggtitle("klpn"))
print(ggtree(phy.klpn.group, aes(color=group))+theme(legend.position="right")+ggtitle("klpn"))
print(ggtree(phy.klpn, layout='circular', aes(color=as.factor(MLST)))+theme(legend.position="right")+ggtitle("klpn"))
print(ggtree(phy.klpn.group, layout='circular', aes(color=group))+theme(legend.position="right")+ggtitle("klpn"))
dev.off()

```
Ok - let's go back to heatmap stuff now for a minute

```{r}

df.klpn=data.frame(meta.klpn$MLST)
rownames(df.klpn)=meta.klpn$label

p1=ggtree(phy.klpn)+geom_tippoint(aes(color=as.factor(Year)))
p2=gheatmap(p1, df.klpn, width=0.2)+theme(legend.position='none')

pdf("~/Dropbox/timplab_data/cpowgs/200707_tree/klpn_tree_mlst_hmap.pdf")
print(p2)
dev.off()
```
