---
title: "Plotting Trees with colorsk"
output: html_notebook
---

```{r setup, include=FALSE}

library(knitr)
library(tidyverse)
library(treeio)
library(readxl)
library(ggtree)
library(rdrop2)
workdir="/atium/Data/Nanopore/cpowgs/trees/200811_timp_tree/"
dboxdir="timplab_data/cpowgs/201103_tree/"
opts_knit$set(root.dir=workdir)

#Pass to bash
Sys.setenv(workdir=workdir)

token=readRDS("~/201104_timpdropbox.rds")
drop_acc(dtoken=token)

```

Ok - the goal for this notebook is to generate plots of the raxml trees  we generated with colored tips, etc from the metadata from Trish

First read the meta info.  I have to pull this back from dropbox - (I used to sync dropbox, but that's gone now)

```{r}

listing=drop_dir(dboxdir, dtoken=token)

listing=listing %>%
  filter(str_detect(listing$name, ".csv$"))

#get files
for(file in listing$path_lower){
  drop_download(file, dtoken=token, overwrite=TRUE)
}


```

Now open those csv files.

```{r, message=FALSE}

meta.klpn=read_csv("klpn.csv") %>%
  select(-X1) %>%
  filter(!(is.na(pilon)&is.na(spades))) %>%
  mutate(label=ifelse(!is.na(pilon), paste0("klpn_pilon_", id), paste0("klpn_spades_", id)))

meta.encl=read_csv("encl.csv") %>%
  select(-X1) %>%
  filter(!(is.na(pilon)&is.na(spades))) %>%
  mutate(label=ifelse(!is.na(pilon), paste0("encl_pilon_", id), paste0("encl_spades_", id)))

meta.esco=read_csv("esco.csv") %>%
  select(-X1) %>%
  filter(!(is.na(pilon)&is.na(spades))) %>%
  mutate(label=ifelse(!is.na(pilon), paste0("esco_pilon_", id), paste0("esco_spades_", id)))

```
Now let's pull in the MLST:

```{r, message=FALSE}

mlst.klpn=read_tsv("klpn_mlst.tsv", col_names=FALSE) %>%
  mutate(label=str_split_fixed(X1, "[/.]",3)[,2]) %>%
  rename(timp_mlst=X3)

meta.klpn=left_join(meta.klpn, mlst.klpn)

mlst.encl=read_tsv("encl_mlst.tsv", col_names=FALSE) %>%
  mutate(label=str_split_fixed(X1, "[/.]",3)[,2]) %>%
  rename(timp_mlst=X3)

meta.encl=left_join(meta.encl, mlst.encl)

mlst.esco=read_tsv("esco_mlst.tsv", col_names=FALSE) %>%
  mutate(label=str_split_fixed(X1, "[/.]",3)[,2]) %>%
  rename(timp_mlst=X3)

meta.esco=left_join(meta.esco, mlst.esco)

```

Point by Trish - can we look for only those MLST types that have more than some threshold.  Let's see what that threshold should be first:

Ok - it looks like 4 might be a natural cutoff

```{r}

filt.meta.klpn=meta.klpn %>%
  group_by(timp_mlst) %>%
  add_tally() %>%
  mutate(filt.mlst=ifelse(n>3, timp_mlst, NA)) %>%
  ungroup()
filt.meta.klpn$filt.mlst[filt.meta.klpn$filt.mlst=="-"]=NA

filt.meta.encl=meta.encl %>%
  group_by(timp_mlst) %>%
  add_tally() %>%
  mutate(filt.mlst=ifelse(n>3, timp_mlst, NA)) %>%
  ungroup()
filt.meta.encl$filt.mlst[filt.meta.encl$filt.mlst=="-"]=NA

filt.meta.esco=meta.esco %>%
  group_by(timp_mlst) %>%
  add_tally() %>%
  mutate(filt.mlst=ifelse(n>3, timp_mlst, NA)) %>%
  ungroup()
filt.meta.esco$filt.mlst[filt.meta.esco$filt.mlst=="-"]=NA

```


Now let's try to get the trees in from the raxml

```{r}

tree.root="/atium/Data/Nanopore/cpowgs/trees/200811_timp_tree"
tree.klpn=read.tree(file.path(tree.root, "raxml_klpn/RAxML_bestTree.testrapid.tree" ))
tbl.klpn=as_tibble(tree.klpn) %>%
  left_join(filt.meta.klpn)
phy.klpn=as.treedata(tbl.klpn)

tree.encl=read.tree(file.path(tree.root, "raxml_encl/RAxML_bestTree.testrapid.tree" ))
tbl.encl=as_tibble(tree.encl) %>%
  left_join(filt.meta.encl)
phy.encl=as.treedata(tbl.encl)

tree.esco=read.tree(file.path(tree.root, "raxml_esco/RAxML_bestTree.testrapid.tree" ))
tbl.esco=as_tibble(tree.esco) %>%
  left_join(filt.meta.esco)
phy.esco=as.treedata(tbl.esco)


```


Now let's plot the tree

```{r}
pdf("plots/trees.pdf")
print(ggtree(phy.klpn)+geom_tippoint(aes(color=as.factor(year)))+theme(legend.position="right")+ggtitle("klpn"))
print(ggtree(phy.esco)+geom_tippoint(aes(color=as.factor(year)))+theme(legend.position="right")+ggtitle("esco"))
print(ggtree(phy.encl)+geom_tippoint(aes(color=as.factor(year)))+theme(legend.position="right")+ggtitle("encl"))
dev.off()
drop_upload("plots/trees.pdf", "timplab_data/cpowgs/201103_tree/", dtoken=token)
```



Ok - replot


```{r}

full_plot<-function(phyobj, meta,label='default') {
  
  ##Ok - first, pull the years out of the phyobj data
  
  hmap.df=data.frame(as.factor(meta$year))
  rownames(hmap.df)=meta$label
  
  p1=ggtree(phyobj)+geom_tippoint(aes(color=as.factor(filt.mlst)))
  p2=gheatmap(p1, hmap.df, width=0.1)+theme(legend.position="right")
  
  plotfile=paste0("plots/", label, "_tree_mlst_hmap.pdf")
  pdf(plotfile)
  print(p2)
  dev.off()
  drop_upload(plotfile, dboxdir, dtoken=token)
}

label_plot<-function(phyobj, meta,label='default') {
  
  ##Ok - first, pull the years out of the phyobj data
  
  hmap.df=data.frame(as.factor(meta$year))
  rownames(hmap.df)=meta$label
  
  p1=ggtree(phyobj)+geom_tippoint(aes(color=as.factor(filt.mlst)))+geom_tiplab()
  p2=gheatmap(p1, hmap.df, width=0.1)+theme(legend.position="right")
  
  plotfile=paste0("plots/", label, "_tree_mlst_hmap.pdf")
  pdf(plotfile, height=20, width=20)
  print(p2)
  dev.off()
  drop_upload(plotfile, dboxdir, dtoken=token)
}


```


```{r}

full_plot(phy.klpn, filt.meta.klpn, label='klpn')
full_plot(phy.esco, filt.meta.esco, label='esco')
full_plot(phy.encl, filt.meta.encl, label='encl')
```

```{r}
label_plot(phy.klpn, filt.meta.klpn, label='trish_klpn')
label_plot(phy.esco, filt.meta.esco, label='trish_esco')
label_plot(phy.encl, filt.meta.encl, label='trish_encl')

```

Write out those filt.meta csvs and push 'em to dbox:

```{r}
write_csv(filt.meta.klpn, "plots/filt.meta.klpn.csv")
drop_upload("plots/filt.meta.klpn.csv", dboxdir, dtoken=token)
write_csv(filt.meta.encl, "plots/filt.meta.encl.csv")
drop_upload("plots/filt.meta.encl.csv", dboxdir, dtoken=token)
write_csv(filt.meta.esco, "plots/filt.meta.esco.csv")
drop_upload("plots/filt.meta.esco.csv", dboxdir, dtoken=token)

```
